Recreation of figures and results in the manuscript

### 2024-04-12 inst/hankin-ejs.Rmd
whitespace

Following is a `sage`-compatible input:

@@

@@

@@

	if(log){
		out <- 0
		for(i in seq_len(nrow(M))){out <- out + f_vec_a(a,r=M[i,1],n=M[i,2],log=TRUE)}
	} else {
		out <- 1
		for(i in seq_len(nrow(M))){out <- out * f_vec_a(a,r=M[i,1],n=M[i,2],log=FALSE)}
	}
	return(out)
}

fapprox <- function(a,r,n,log=FALSE){
	if(log){
		return(log(B-1)-B*(r+1)/n)
	} else {
		return(B-1 - (r+1)*exp(B)/n)
	}
}
```

@@

Now verify `fdash()`:

@@

## Maximum likelihood estimation

@@

@@

Now verification

@@

@@

Pure first

```{r dopurefurst}
Lpure <- f_vec_arn(a,rank[wp]-1,class_size[wp]-1,log=TRUE)
Lpure <- Lpure - max(Lpure,na.rm=TRUE)
plot(a,Lpure,ylab='pure strength')
optimize(function(a){f_vec_arn(a,rank[wp]-1,class_size[wp]-1,log=TRUE)},c(0.3,0.6),maximum=TRUE)
optimize(function(a){f_vec_arn(a,rank[wa]-1,class_size[wa]-1,log=TRUE)},c(0.3,0.6),maximum=TRUE)
```

@@

@@

@@

>>>>

First we use Stirling's approximation $\log(n!)\sim
(n+1/2)\log(n)-n+\log(2\pi)/2$ to simplify equation 4 of the manuscript
Following is a `sage`-compatible input [the $\sqrt{2\pi}$ bit
cancels]:

@@

where $K$ is a constant.

@@

Below I give R idiom for the ikelihood expression given by equation 4
in the manuscript.  Some inline comments are included.

@@

	if(log){
		out <- 0
		for(i in seq_len(nrow(M))){out <- out + f_vec_a(a,r=M[i,1],n=M[i,2],log=TRUE)}
	} else {
		out <- 1
		for(i in seq_len(nrow(M))){out <- out * f_vec_a(a,r=M[i,1],n=M[i,2],log=FALSE)}
	}
	return(out)
}
```

```{r definefapprox}
fapprox <- function(a,r,n,log=FALSE){
	if(log){
		return(log(B-1)-B*(r+1)/n)
	} else {
		return(B-1 - (r+1)*exp(B)/n)
	}
}
```

Above, `fapprox()` gives the asymptotic expression (5) in the
manuscript.

@@

Notation is $\mathcal{L}$ for likelihood and
$\mathcal{S}=\log\mathcal{L}$ for log-likelihood (support).  Above we
see `fdash()` implements either $\frac{\partial\mathcal{S}}{\partial
a}$ or $\frac{\partial\mathcal{L}}{\partial a}$ [depending on the
value of argument `log`.  We see that

$$
\frac{\partial\mathcal{S}}{\partial a}=
B^2\left(\frac{1}{B-1}-\psi(B+n)+\psi(B+n-r-1)\right)$$

and $\frac{\partial\mathcal{L}}{\partial a}$ is of course just
$\frac{\mathcal{L}\partial\mathcal{S}}{\partial a}$.  Now, numerical
verification of `fdash()`:

@@

Above we see very close agreemen between numerical and analytical
results, for both `log=TRUE` and `log=FALSE`.

## Maximum likelihood estimation {-}

@@

Above we see very close agreement between the two methods.  We can get
a feel for the accuracy of the asymptotic result numerically using
similar methods:

@@

Now another verification, this time comparing the R implementation
against two different direct numerical implementations of equation 4
of the manuscript:

@@

Above we see close agreement.

@@

Following lines show optimization of the _marginal_ Bradley Terry
strength for pure and applied mathematics:

```{r dopurefurst}
Lpure <- f_vec_arn(a,rank[wp]-1,class_size[wp]-1,log=TRUE)
Lpure <- Lpure - max(Lpure,na.rm=TRUE)
plot(a,Lpure,ylab='pure strength')
optimize(function(a){f_vec_arn(a,rank[wp]-1,class_size[wp]-1,log=TRUE)},c(0.1,0.9),maximum=TRUE)
optimize(function(a){f_vec_arn(a,rank[wa]-1,class_size[wa]-1,log=TRUE)},c(0.1,0.9),maximum=TRUE)
```

[the evaluates above do not appear in the manuscript] Now we define a
likelihood function for the two subjects jointly:

@@

Following code shows _joint_ maximization of likelihood:

@@

Following code shows _constrained_ maximization of likelihood
[constrained, that is, to follow the diagonal null line]:

@@

Extra support is thus $-86.9449 + 89.3732=2.4283$ (to 4 d.p.)  Now
calculate the asymptotic $p$-value obtained from Wilks's theorem:
